Software Developers and Computers as Social Actors

### 2016-03-21 paper/draft1.tex
Renamed "sig..." to draft1.tex.  Set up CI.

% DOI
%\doi{10.475/123_4}

% ISBN
%\isbn{123-4567-24-567/08/06}

%Conference
%\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

\acmPrice{\$15,000}

%
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
\CopyrightYear{2016} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Do Software Developers Love Their Computers?}
%\subtitle{A Nass Replication}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\maketitle
\begin{abstract}
\begin{comment}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings. It is an {\em alternate} style which produces
a {\em tighter-looking} paper and was designed in response to
concerns expressed, by authors, over page-budgets.
It complements the document \textit{Author's (Alternate) Guide to
Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
This source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file. A compiled PDF version
is available on the web page to help you with the
`look and feel'.
\end{comment}
\end{abstract}

@@

%
% End generated code
%

@@

% We no longer use \terms command
%\terms{Theory}

\keywords{ACM proceedings; \LaTeX; text tagging}

\section{Introduction}

\section{Related Work}
\begin{comment}
In research interviews open, frank speech is essential to the process. Interviews are plagued by biases inherent in social interactions. Such biases include the normative-response bias (people not wanting to admit to answers that are counter to the perceived norm), interviewer-based bias (people answering based on the perceived preference of the interviewer), and politeness norms (people modifying answers to avoid offending). Computer systems were introduced to mitigate these known social-biases by providing anonymity. However, through a series of studies, Nass et al. documented how individuals anthropomorphize technology and exhibit attitudes previously thought to be reserved for person to person interaction \cite{nass1999people}\cite{reeves1996people}. Contrary to what was believed at the time, the results suggested that humans unintentionally, unknowingly apply social rules to computer systems.   
This foundational work sparked further studies to replicate, validate, and extend this assertion. Since the initial set of studies were conducted entirely in the United States, additional research validated the findings across other cultures \cite{katagiri2001cross}. Furthermore, the existence of computers as social entities has been studied to better understand racial biases in interviews\cite{krysan2003race}, how the quality of interaction affects expressive systems \cite{vidyarthi2011sympathetic}, and even how to leverage social cues to more effectively sell products online \cite{wang2007can}. To our knowledge, no studies currently exist which attempt to replicate these effects with software developers.   
In order to create better software, it is important to study the attitudes of software developers and how they view their work environments. Given software developers’ advanced understanding of how computers work, one would expect the attribution of social qualities to computers to be minimal or not present. At the same time, it has been shown that humans with accurate mental models are more effective in maintaining teammate bonds with automated computer systems \cite{wilkison2008effects}. Given this premise, we are interested in exploring the following questions:

@@

{\color{red} Draft, do not submit}\\
Researchers have been investigating the phenomena of human users considering computers as social entities. There have been studies that focus on the human application of social norms and expectations to computers in a seamless manner. Nass et al. performed a series of experiments that pointed out that social stereotypes attributed to gender, ethnicity, and group loyalty, along with deeply ingrained social habits and behaviors like, reciprocity and reciprocal self-disclosure, are extended to computers when interacting with them~\cite{Nass2000machines,nass1999people}. Moreover, prior studies document that this behavior is not only limited to computers, human users anthropomorphize other technological mediums like television~\cite{reeves1996people}.

These findings engendered a number of studies that extended on this implication by replicating, validating, and exploring the possible factors behind this kind of responses. These existing Computers as Social Actors (CASA) studies featured an implicit bias - although the studies were conducted with participants exclusively from the United States, the findings were generalized to all cultures. Katagiri and collegues performed a comprehensive study to determine whether social rules derived from different cultures affect the human users responses to computers by observing participants from the United States and Japan~\cite{katagiri2001cross}. Furthermore, the existence of computers as social entities has been studied to better understand racial biases in interviews\cite{krysan2003race}, how the quality of interaction affects expressive systems \cite{vidyarthi2011sympathetic}, and even how to leverage social cues to more effectively sell products online \cite{wang2007can}. In contrast, our work explores the effect of viewing computers as a human-like social entity on the very specific demographic of software developers and attempt to determine the level of politeness towards their machines.

Software developers maintain a collaborative environment with their development machines to achieve their goals. Given software developers' advanced understanding of how computers work, one would expect the attribution of social qualities to computers to be minimal or not present, if the the sociability of a computer system is modified by the quality of an individual's mental model of a computer system. Research has shown that human-computer team performance is higher when human users have accurate mental models~\cite{wilkison2008effects}. Further studies by Bickmore and Picard have documented that human users of a relational computer agent did not only respect, trust, and like the agent, but also wanted to continue working with the agent even after the end of the study~\cite{Bickmore:2005:EML:1067860.1067867}. In this paper, we investigate whether such long term relationship between a software developer and her computer has any significant effect on the social norm of politeness witnessed by Nass et al in their original study~\cite{nass1999people} by contrasting their responses to computer with that of non-software developers .
\\{\color{red} Draft, do not submit}
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%
% That's all folks!
\end{document}

>>>>

% DOI
\doi{}

% ISBN
\isbn{}

%Conference
%\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}
%\acmPrice{\$15,000}

%
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2016} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Software Developers and Computers as Social Actors}
%\subtitle{A Nass Replication}
%

@@

\maketitle
\begin{abstract}
\begin{itemize}
    \item Standard abstract, defines what we were able to achieve.
\end{itemize}
\end{abstract}

@@

@@

\keywords{software engineering; computers as social actors}

\section{Introduction}
\begin{itemize}
    \item Define what Nass' study was.
    \item Define what is novel about our study in comparison to Nass' original study.
    \item Introduce who our subjects are and the basics of what our methodology was.
\end{itemize}

\section{Related Work}

\begin{comment}
In research interviews open, frank speech is essential to the process. Interviews are plagued by biases inherent in social interactions. Such biases include the normative-response bias (people not wanting to admit to answers that are counter to the perceived norm), interviewer-based bias (people answering based on the perceived preference of the interviewer), and politeness norms (people modifying answers to avoid offending). Computer systems were introduced to mitigate these known social-biases by providing anonymity. However, through a series of studies, Nass et al. documented how individuals anthropomorphize technology and exhibit attitudes previously thought to be reserved for person to person interaction \cite{nass1999people}\cite{reeves1996people}. Contrary to what was believed at the time, the results suggested that humans unintentionally, unknowingly apply social rules to computer systems.   
This foundational work sparked further studies to replicate, validate, and extend this assertion. Since the initial set of studies were conducted entirely in the United States, additional research validated the findings across other cultures \cite{katagiri2001cross}. Furthermore, the existence of computers as social entities has been studied to better understand racial biases in interviews\cite{krysan2003race}, how the quality of interaction affects expressive systems \cite{vidyarthi2011sympathetic}, and even how to leverage social cues to more effectively sell products online \cite{wang2007can}. To our knowledge, no studies currently exist which attempt to replicate these effects with software developers.   
In order to create better software, it is important to study the attitudes of software developers and how they view their work environments. Given software developers’ advanced understanding of how computers work, one would expect the attribution of social qualities to computers to be minimal or not present. At the same time, it has been shown that humans with accurate mental models are more effective in maintaining teammate bonds with automated computer systems \cite{wilkison2008effects}. Given this premise, we are interested in exploring the following questions:

@@

Previous research demonstrates that computers are often treated as social entities despite their inanimate nature. There have been studies that focus on the human application of social norms and expectations to computers in a seamless manner. Nass et al. performed a series of experiments that pointed out that social stereotypes attributed to gender, ethnicity, and group loyalty, along with deeply ingrained social habits and behaviors like, reciprocity and reciprocal self-disclosure, are extended to computers when interacting with them~\cite{Nass2000machines,nass1999people}. Moreover, prior studies documented that this behavior is not only limited to computers, human users anthropomorphize other technological mediums like television~\cite{reeves1996people}.

\hfill

These findings engendered a number of studies that extended on this implication by replicating, validating, and exploring the possible factors behind these kind of responses. The existing Computers as Social Actors (CASA) studies featured an implicit bias - although the studies were conducted with participants exclusively from the United States, the findings were generalized to all cultures. Katagiri and colleagues performed a comprehensive study to determine whether social rules derived from different cultures affect human users responses to computers by observing participants from the United States and Japan~\cite{katagiri2001cross}. Furthermore, researchers studied the existence of computers as social entities to better understand racial biases in interviews\cite{krysan2003race}, how the quality of interaction affects expressive systems \cite{vidyarthi2011sympathetic}, and even how to leverage social cues to more effectively sell products online \cite{wang2007can}.

\hfill

Software developers maintain a collaborative environment with their development machines to achieve their goals. Given software developers' advanced understanding of how computers work, one would expect the attribution of social qualities to computers to be minimal or not present.  This type of attribution would be expected, if the the sociability of a computer system is modified by the quality of an individual's mental model of a computer system. Research has shown that human-computer team performance is higher when human users have accurate mental models~\cite{wilkison2008effects}. Further studies by Bickmore and Picard have documented the long-term relation establishment and maintenance between a human-computer team by observing that human users of a relational computer agent did not only respect, trust, and like the agent, but also wanted to continue working with the agent even after the end of the study~\cite{Bickmore:2005:EML:1067860.1067867}.

\hfill

In this paper, we investigate whether a long term relationship between a software developer and her computer has any significant effect on the social norm of politeness witnessed by Nass et al in their original study~\cite{nass1999people} by contrasting their responses to computers with that of non-software developers.

Our work explores the effect of viewing computers as a human-like social entity on the very specific demographic of software developers and attempts to determine the level of politeness software developers exhibit toward their machines.

\section{Methods}

\subsection{Participants}
Here we will detail the composition of our study participants including:
\begin{itemize}
    \item The number in each group
    \item Programming experience breakdown
    \item Gender breakdown (potentially)
\end{itemize}

\subsection{Procedure}
Generally, here we will detail the procedure by which the study was conducted.

\begin{itemize}
    \item Screenshots of the website which participants visited.
    \item Detailed information about the experimental environment: 
    \begin{itemize}
        \item Where the study took place.
        \item What machines people used.
        \item How long the study took.
        \item Explaination of the function of the website used.
    \end{itemize}
\end{itemize}

\section{Results}

\begin{itemize}
    \item Overall description of what our results were.
\end{itemize}

\subsection{Analysis}

\begin{itemize}
    \item Graph describing the differences between CSC group and Psychology group, along with discussion of that graph.
    \item Table describing differences between CSC groups and Psych group, with discussion.
    \item Appropriate statistical tests to describe potential inferences that can be made from the results.
\end{itemize}

\section{Discussion}

\begin{itemize}
    \item Overall discussion of the importance or lack of importance of our analysis and results.
\end{itemize}

\subsection{Threats to Validity}

\subsubsection{Internal Validity}

\begin{itemize}
    \item Describe the mistakes we made.
\end{itemize}

\subsubsection{External Validity}

\begin{itemize}
    \item Describe whether our results are significant enough to generalize to future studies.
\end{itemize}

\subsection{Future Work}

\begin{itemize}
    \item Describe any potential studies that we believe can build on our own.
\end{itemize}

\section{Conclusion}
Although I'm not a super big fan of conclusions, it makes sense to have something short here I'd think.

\section*{Acknowledgements}
We would like to thank...

\bibliographystyle{abbrv}
\bibliography{sigproc}

\appendix
\section{Website and Survey Transcripts}

\end{document}
